---
title: "Clinical Benefit Fairness"
author: "Gaylen Fronk"
date: "`r lubridate::today()`"
output: 
  html_document:
    toc: true 
    toc_depth: 4
format:
  html:
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

## Set Up Environment

```{r}
study <- "match"
version <- "v6"
cv <- "nested"
y_col_name <- "pp_hybrid_wk4_outcome"
```

Function conflicts
```{r}
#| message: false
#| warning: false

# handle conflicts
options(conflicts.policy = "depends.ok")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_ml.R?raw=true")
tidymodels_conflictRules()
```

Packages, functions, and paths
```{r}
#| message: false
#| warning: false

library(tidyverse)
library(tidymodels)

devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/chtc/static_files/fun_chtc.R?raw=true")



path_processed <- "R:/studydata/match/processed_data"
path_models <- str_c("R:/studydata/match/models/", y_col_name)
```

## Prepare data

Read in demographics

```{r}

d_dem <- read_csv(file.path(path_processed, "match_cln.csv"), 
              show_col_types = FALSE) |> 
  select(subid, gender_ehr, race_ehr, ethnicity_ehr, income_order) |> 
  mutate(sex = if_else(gender_ehr == "female", "Female", "Male"),
         race_ethnicity = if_else(race_ehr == "white" & ethnicity_ehr == "non_hispanic",
                                  "White/Non-Hispanic", "Not White"),
         income = case_match(
           income_order,
           c("01_less_than_10000", "02_btwn_10000_19999", 
             "03_btwn_20000_24999") ~ "Below Poverty Line",
           c("04_btwn_25000_34999", "05_btwn_35000_49999", 
             "06_btwn_50000_74999", "07_more_than_75000") ~ "Above Poverty Line",
           NA_character_ ~ "Missing")) |> 
  select(-income_order, -ends_with("ehr")) |> 
  glimpse()
```

Read in aim 2 dataset

```{r}

d_ben <- read_csv(file.path(path_models, 
                        str_c("aim_2_", version, "_", y_col_name, ".csv")),
              show_col_types = FALSE) |> 
  mutate(outcome_rct_wk4_num = if_else(outcome_rct_wk4 == "abstinent", 1, 0),
         outcome_rct_wk12_num = if_else(outcome_rct_wk12 == "abstinent", 1, 0),
         outcome_rct_wk26_num = if_else(outcome_rct_wk26 == "abstinent", 1, 0),
         tx_worst = case_when(
           prob_patch < prob_combo_nrt & prob_patch < prob_varenicline ~ "patch",
           prob_combo_nrt < prob_patch & prob_combo_nrt < prob_varenicline ~ "combo_nrt",
           prob_varenicline < prob_patch & prob_varenicline < prob_combo_nrt ~ "varenicline",
           TRUE ~ NA_character_),
         tx_second = case_when(
           tx_worst == "patch" & tx_best == "varenicline" ~ "combo_nrt",
           tx_worst == "patch" & tx_best == "combo_nrt" ~ "varenicline",
           tx_worst == "varenicline" & tx_best == "patch" ~ "combo_nrt",
           tx_worst == "varenicline" & tx_best == "combo_nrt" ~ "patch",
           tx_worst == "combo_nrt" & tx_best == "varenicline" ~ "patch",
           tx_worst == "combo_nrt" & tx_best == "patch" ~ "varenicline",
           TRUE ~ NA_character_)) |> 
  select(subid, starts_with("tx_"), starts_with("prob_"),
         outcome_rct_wk4_num, outcome_rct_wk12_num, outcome_rct_wk26_num) |> 
  pivot_longer(
    cols = c(outcome_rct_wk4_num, outcome_rct_wk12_num, outcome_rct_wk26_num),
    names_to = "week",
    names_pattern = "(?<=outcome_rct_wk)(.+)(?=_num)",
    values_to = "outcome_rct_num"
  ) |> 
  mutate(c_tx_match = if_else(tx_match == TRUE, 0.5, -0.5), # center
         week = as.numeric(week))

glimpse(d_ben)

```

Join
```{r}
d <- d_ben |> 
  left_join(d_dem, by = "subid") |> 
  glimpse()

```

## Analyses

Centered variables
```{r}
d <- d |> 
  mutate(
    c_sex_male = if_else(sex == "Male", 0, 1),
    c_sex_female = if_else(sex == "Male", 1, 0),
    c_race_ethn_white_non_his = if_else(race_ethnicity == "White/Non-Hispanic", 0, 1),
    c_race_ethn_not_white = if_else(race_ethnicity == "White/Non-Hispanic", 1, 0),
    c_income_above = case_match(
      income,
      "Below Poverty Line" ~ 1,
      "Above Poverty Line" ~ 0,
      "Missing" ~ NA_real_
    ),
    c_income_below = case_match(
      income,
      "Below Poverty Line" ~ 0,
      "Above Poverty Line" ~ 1,
      "Missing" ~ NA_real_
    )
  ) |> 
  glimpse()
```

Filter to week 4
```{r}
d_wk4 <- d |> 
  filter(week == 4)

```

### Race/ethnicity

Simple effect of White/Non-Hispanic at week 4
```{r}
white_4 <- glm(outcome_rct_num ~ c_tx_match * c_race_ethn_white_non_his,
                data = d_wk4,
                family = binomial(link = "logit"))

summary(white_4)
exp(coef(white_4)["c_tx_match"])
```

Simple effect of Non-White at week 4
```{r}
nonwhite_4 <- glm(outcome_rct_num ~ c_tx_match * c_race_ethn_not_white,
                data = d_wk4,
                family = binomial(link = "logit"))

summary(nonwhite_4)
exp(coef(nonwhite_4)["c_tx_match"])
```

Overall effect OR: 1.38
White/non-Hispanic OR: 1.208
Non-White OR: 2.164

```{r}
fig_clin_ben_race <- d_wk4 |> 
  mutate(race_ethn_cat = factor(c_race_ethn_not_white,
                           levels = c(0, 1),
                           labels = c("Not White", "White/Non-Hispanic")),
         tx_match = factor(tx_match, 
                           levels = c(TRUE, FALSE),
                           labels = c("Best Treatment", "Other Treatment"))) |> 
  group_by(race_ethn_cat, tx_match) |> 
  summarize(mean_outcome = mean(outcome_rct_num),
            sd_outcome = sd(outcome_rct_num),
            n_group = n()) |> 
  mutate(se_outcome = (sd_outcome / sqrt(n_group))) |> 
  ggplot(aes(x = race_ethn_cat, y = mean_outcome, fill = tx_match)) +
  geom_col(position = "dodge") +
  geom_errorbar(aes(ymin = mean_outcome - se_outcome, 
                    ymax = mean_outcome + se_outcome,
                    x = race_ethn_cat),
                position = position_dodge(width = 0.9),
                width = 0.5) +
  scale_y_continuous(limits = c(0, 0.55)) +
  labs(
    x = "Race/Ethnicity",
    y = "Mean Abstinence Rate",
    fill = "Treatment Matching"
  )

fig_clin_ben_race
```

### Sex

Simple effect of Male at week 4
```{r}
male_4 <- glm(outcome_rct_num ~ c_tx_match * c_sex_male,
                data = d_wk4,
                family = binomial(link = "logit"))

summary(male_4)
exp(coef(male_4)["c_tx_match"])
```

Simple effect of Female at week 4
```{r}
female_4 <- glm(outcome_rct_num ~ c_tx_match * c_sex_female,
                data = d_wk4,
                family = binomial(link = "logit"))

summary(female_4)
exp(coef(female_4)["c_tx_match"])
```

Overall effect OR: 1.38
Female OR: 1.372
Male OR: 1.395

```{r}
fig_clin_ben_sex <- d_wk4 |> 
  mutate(sex_cat = factor(c_sex_female,
                           levels = c(0, 1),
                           labels = c("Female", "Male")),
         tx_match = factor(tx_match, 
                           levels = c(TRUE, FALSE),
                           labels = c("Best Treatment", "Other Treatment"))) |> 
  group_by(sex_cat, tx_match) |> 
  summarize(mean_outcome = mean(outcome_rct_num),
            sd_outcome = sd(outcome_rct_num),
            n_group = n()) |> 
  mutate(se_outcome = (sd_outcome / sqrt(n_group))) |> 
  ggplot(aes(x = sex_cat, y = mean_outcome, fill = tx_match)) +
  geom_col(position = "dodge") +
  geom_errorbar(aes(ymin = mean_outcome - se_outcome, 
                    ymax = mean_outcome + se_outcome,
                    x = sex_cat),
                position = position_dodge(width = 0.9),
                width = 0.5) +
  scale_y_continuous(limits = c(0, 0.55)) +
  labs(
    x = "Sex",
    y = "Mean Abstinence Rate",
    fill = "Treatment Matching"
  )

fig_clin_ben_sex
```

### Income

Need to remove subset of participants who did not provide income data.
```{r}
d_wk4_income <- d_wk4 |> 
  filter(income != "Missing") |> 
  mutate(c_income = if_else(income == "Above Poverty Line", -0.5, 0.5))
```

Overall week 4 (centered)
```{r}
centered_4 <- glm(outcome_rct_num ~ c_tx_match * c_income,
                data = d_wk4_income,
                family = binomial(link = "logit"))
summary(centered_4)
exp(coef(centered_4)["c_tx_match"])
```

Simple effect of higher income at week 4
```{r}
above_4 <- glm(outcome_rct_num ~ c_tx_match * c_income_above,
                data = d_wk4_income,
                family = binomial(link = "logit"))

summary(above_4)
exp(coef(above_4)["c_tx_match"])
```

Simple effect of low income at week 4
```{r}
below_4 <- glm(outcome_rct_num ~ c_tx_match * c_income_below,
                data = d_wk4_income,
                family = binomial(link = "logit"))

summary(below_4)
exp(coef(below_4)["c_tx_match"])
```

Overall effect OR: 1.482
High income OR: 1.437
Low income OR: 1.529

```{r}
fig_clin_ben_income <- d_wk4_income |> 
  mutate(income_cat = factor(c_income_above,
                           levels = c(1, 0),
                           labels = c("Below Poverty Line", "Above Poverty Line")),
         tx_match = factor(tx_match, 
                           levels = c(TRUE, FALSE),
                           labels = c("Best Treatment", "Other Treatment"))) |> 
  group_by(income_cat, tx_match) |> 
  summarize(mean_outcome = mean(outcome_rct_num),
            sd_outcome = sd(outcome_rct_num),
            n_group = n()) |> 
  mutate(se_outcome = (sd_outcome / sqrt(n_group))) |> 
  ggplot(aes(x = income_cat, y = mean_outcome, fill = tx_match)) +
  geom_col(position = "dodge") +
  geom_errorbar(aes(ymin = mean_outcome - se_outcome, 
                    ymax = mean_outcome + se_outcome,
                    x = income_cat),
                position = position_dodge(width = 0.9),
                width = 0.5) +
  scale_y_continuous(limits = c(0, 0.55)) +
  labs(
    x = "Income",
    y = "Mean Abstinence Rate",
    fill = "Treatment Matching"
  )

fig_clin_ben_income
```




