---
title: "Evaluation of Clinical Benefit: Week 4 Model"
author: "Gaylen Fronk"
date: "`r lubridate::today()`"
number-sections: true
output: 
  html_document:
    toc: true 
    toc_depth: 4
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 4
editor_options: 
  chunk_output_type: console
---

```{r set_variables}
study <- "match"
version <- "v6"
y_col_name <- "pp_hybrid_wk4_outcome"
```

```{r, packages_script}
#| message: false
#| warning: false

# packages for script
library(tidyverse)
library(blme)
library(lme4)

devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_eda.R?raw=true")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/chtc/static_files/fun_chtc.R?raw=true")
theme_set(theme_classic()) 
```

```{r, packages_workflow}
#| message: false
#| warning: false

# handle conflicts
options(conflicts.policy = "depends.ok")
```

```{r, absolute paths}

# absolute paths
switch (Sys.info()[['sysname']],
        # PC paths
        Windows = {
          path_models <- str_c("P:/models/", y_col_name)
          path_data <- "P:/processed_data"},
        
        # IOS paths
        Darwin = {
          path_models <- str_c("/Volumes/models/",
                               y_col_name)
          path_data <- "/Volumes/processed_data"},
        
        # Linux paths
        Linux = {
          path_models <- str_c("~/mnt/models/",
                               y_col_name)
          path_data <- "~/mnt/processed_data"}
)
```

```{r defaults}

# chunk defaults

options(tibble.width = Inf)
options(tibble.print_max = Inf)
```

Read in aim 2 dataset
```{r}

d <- read_csv(file.path(path_models, 
                        str_c("aim_2_", version, "_", y_col_name, ".csv")),
              show_col_types = FALSE) |> 
  mutate(outcome_rct_wk4_num = if_else(outcome_rct_wk4 == "abstinent", 1, 0),
         outcome_rct_wk12_num = if_else(outcome_rct_wk12 == "abstinent", 1, 0),
         outcome_rct_wk26_num = if_else(outcome_rct_wk26 == "abstinent", 1, 0),
         tx_worst = case_when(
           prob_patch < prob_combo_nrt & prob_patch < prob_varenicline ~ "patch",
           prob_combo_nrt < prob_patch & prob_combo_nrt < prob_varenicline ~ "combo_nrt",
           prob_varenicline < prob_patch & prob_varenicline < prob_combo_nrt ~ "varenicline",
           TRUE ~ NA_character_),
         tx_second = case_when(
           tx_worst == "patch" & tx_best == "varenicline" ~ "combo_nrt",
           tx_worst == "patch" & tx_best == "combo_nrt" ~ "varenicline",
           tx_worst == "varenicline" & tx_best == "patch" ~ "combo_nrt",
           tx_worst == "varenicline" & tx_best == "combo_nrt" ~ "patch",
           tx_worst == "combo_nrt" & tx_best == "varenicline" ~ "patch",
           tx_worst == "combo_nrt" & tx_best == "patch" ~ "varenicline",
           TRUE ~ NA_character_)) |> 
  mutate(tx_rank = case_when(
    tx_rct == tx_best ~ "first",
    tx_rct == tx_second ~ "second",
    tx_rct == tx_worst ~ "third",
    TRUE ~ NA_character_)) |> 
  select(subid, starts_with("tx_"), starts_with("prob_"),
         outcome_rct_wk4_num, outcome_rct_wk12_num, outcome_rct_wk26_num) |> 
  pivot_longer(
    cols = c(outcome_rct_wk4_num, outcome_rct_wk12_num, outcome_rct_wk26_num),
    names_to = "week",
    names_pattern = "(?<=outcome_rct_wk)(.+)(?=_num)",
    values_to = "outcome_rct_num"
  ) |> 
  mutate(c_tx_match = if_else(tx_match == TRUE, 0.5, -0.5), # center
         week = as.numeric(week))

glimpse(d)

```

Read in start-medication data and join
```{r}
d_med <- read_csv(file.path(path_data, "match_used_meds_cln.csv"),
                  show_col_types = FALSE)

glimpse(d_med)

d <- left_join(d, d_med, join_by("subid", "tx_rct" == "treatment"))

glimpse(d)
```

## Analyses

```{r}
d_wk4 <- d |> 
  filter(week == 4)

d_wk12 <- d |> 
  filter(week == 12)

d_wk26 <- d |> 
  filter(week == 26)
```


### Primary

4 weeks
```{r}
model_4wk <- glm(outcome_rct_num ~ c_tx_match, 
                 data = d_wk4,
                 family = binomial(link = "logit"))

summary(model_4wk)
```

12 weeks
```{r}
model_12wk <- glm(outcome_rct_num ~ c_tx_match, 
                  data = d_wk12,
                  family = binomial(link = "logit"))

summary(model_12wk)
```

26 weeks
```{r}
model_26wk <- glm(outcome_rct_num ~ c_tx_match, 
                  data = d_wk26,
                  family = binomial(link = "logit"))

summary(model_26wk)
```

### Supplemental analyses: Started medication subset

```{r}
d_meds_only <- d |> 
  filter(used_any_meds == "started_meds")

dim(d_meds_only)

d_meds_wk4 <- d_meds_only |> 
  filter(week == 4)

d_meds_wk12 <- d_meds_only |> 
  filter(week == 12)

d_meds_wk26 <- d_meds_only |> 
  filter(week == 26)
```

4 weeks
```{r}
model_4wk_meds <- glm(outcome_rct_num ~ c_tx_match, 
                      data = d_meds_wk4,
                      family = binomial(link = "logit"))

summary(model_4wk_meds)
```
Still significant, slightly smaller estimate

12 weeks
```{r}
model_12wk_meds <- glm(outcome_rct_num ~ c_tx_match, 
                       data = d_meds_wk12,
                       family = binomial(link = "logit"))

summary(model_12wk_meds)
```
Still non-significant

26 weeks
```{r}
model_26wk_meds <- glm(outcome_rct_num ~ c_tx_match, 
                       data = d_meds_wk26,
                       family = binomial(link = "logit"))

summary(model_26wk_meds)
```
Still non-significant

### Supplemental analyses: Treatment rank

```{r}
d <- d |> 
  mutate(tx_rank_1v23 = case_match(tx_rank,
                                   "first" ~ 2/3,
                                   c("second", "third") ~ -1/3),
         tx_rank_2v3 = case_match(tx_rank, 
                                  "first" ~ 0,
                                  "third" ~ -.5,
                                  "second" ~ .5))

d_wk4 <- d |> 
  filter(week == 4)

d_wk12 <- d |> 
  filter(week == 12)

d_wk26 <- d |> 
  filter(week == 26)
```

4 weeks
```{r}
model_4wk_rank <- glm(outcome_rct_num ~ tx_rank_1v23 + tx_rank_2v3, 
                 data = d_wk4,
                 family = binomial(link = "logit"))

summary(model_4wk_rank)
```

12 weeks
```{r}
model_12wk_rank <- glm(outcome_rct_num ~ tx_rank_1v23 + tx_rank_2v3, 
                  data = d_wk12,
                  family = binomial(link = "logit"))

summary(model_12wk_rank)
```

26 weeks
```{r}
model_26wk_rank <- glm(outcome_rct_num ~ tx_rank_1v23 + tx_rank_2v3, 
                  data = d_wk26,
                  family = binomial(link = "logit"))

summary(model_26wk_rank)
```

## Figures

Primary figure
```{r}
#| label: fig-clin-ben-wk4
#| fig-cap: "Benefit of treatment matching. Bars represent mean observed abstinence (from original trial) for individuals who did and did not receive their model-predicted best treatment, over time. Error bars indicate standard errors."

fig_clin_ben_wk4 <- d |> 
  mutate(week_cat = factor(week,
                           levels = c(4, 12, 26),
                           labels = c("Week 4", "Week 12", "Week 26")),
         tx_match = factor(tx_match, 
                           levels = c(TRUE, FALSE),
                           labels = c("Best Treatment", "Other Treatment"))) |> 
  group_by(week_cat, tx_match) |> 
  summarize(mean_outcome = mean(outcome_rct_num),
            sd_outcome = sd(outcome_rct_num),
            n_group = n()) |> 
  mutate(se_outcome = (sd_outcome / sqrt(n_group))) |> 
  ggplot(aes(x = week_cat, y = mean_outcome, fill = tx_match)) +
  geom_col(position = "dodge") +
  geom_errorbar(aes(ymin = mean_outcome - se_outcome, 
                    ymax = mean_outcome + se_outcome,
                    x = week_cat),
                position = position_dodge(width = 0.9),
                width = 0.5) +
  scale_y_continuous(limits = c(0, 0.45)) +
  labs(
    x = "Week On-Study",
    y = "Mean Abstinence Rate",
    fill = "Treatment Matching"
  )

fig_clin_ben_wk4
```

Primary figure (replicated for supplement)
```{r}
#| label: fig-clin-ben-wk4-supp
#| fig-cap: "(Supplemental; reproduced from main text) Benefit of treatment matching from 4-week prediction model. Bars represent mean observed abstinence (from original trial) for individuals who did and did not receive their model-predicted best treatment, over time. Error bars indicate standard errors."

fig_clin_ben_wk4_supp <- d |> 
  mutate(week_cat = factor(week,
                           levels = c(4, 12, 26),
                           labels = c("Week 4", "Week 12", "Week 26")),
         tx_match = factor(tx_match, 
                           levels = c(TRUE, FALSE),
                           labels = c("Best Treatment", "Other Treatment"))) |> 
  group_by(week_cat, tx_match) |> 
  summarize(mean_outcome = mean(outcome_rct_num),
            sd_outcome = sd(outcome_rct_num),
            n_group = n()) |> 
  mutate(se_outcome = (sd_outcome / sqrt(n_group))) |> 
  ggplot(aes(x = week_cat, y = mean_outcome, fill = tx_match)) +
  geom_col(position = "dodge") +
  geom_errorbar(aes(ymin = mean_outcome - se_outcome, 
                    ymax = mean_outcome + se_outcome,
                    x = week_cat),
                position = position_dodge(width = 0.9),
                width = 0.5) +
  scale_y_continuous(limits = c(0, 0.45)) +
  labs(
    x = "Week On-Study",
    y = "Mean Abstinence Rate",
    fill = "Treatment Matching"
  )

fig_clin_ben_wk4_supp
```

Supplemental figure: Started medication subset
```{r}
#| label: fig-med-subset-wk4-supp
#| fig-cap: "Benefit of treatment matching among individuals who reported any medication use over the first 4 weeks. Bars represent mean observed abstinence (from original trial) for individuals who did and did not receive their model-predicted best treatment, over time. Error bars indicate standard errors."

fig_med_subset_wk4_supp <- d_meds_only |> 
  mutate(week_cat = factor(week,
                           levels = c(4, 12, 26),
                           labels = c("Week 4", "Week 12", "Week 26")),
         tx_match = factor(tx_match, 
                           levels = c(TRUE, FALSE),
                           labels = c("Best Treatment", "Other Treatment"))) |> 
  group_by(week_cat, tx_match) |> 
  summarize(mean_outcome = mean(outcome_rct_num),
            sd_outcome = sd(outcome_rct_num),
            n_group = n()) |> 
  mutate(se_outcome = (sd_outcome / sqrt(n_group))) |> 
  ggplot(aes(x = week_cat, y = mean_outcome, fill = tx_match)) +
  geom_col(position = "dodge") +
  geom_errorbar(aes(ymin = mean_outcome - se_outcome, 
                    ymax = mean_outcome + se_outcome,
                    x = week_cat),
                position = position_dodge(width = 0.9),
                width = 0.5) +
  scale_y_continuous(limits = c(0, 0.45)) +
  labs(
    x = "Week On-Study",
    y = "Mean Abstinence Rate",
    fill = "Treatment Matching"
  )

fig_med_subset_wk4_supp
```

Supplemental figure: Treatment rank
```{r}
#| label: fig-rank-wk4-supp
#| fig-cap: "Benefit of treatment matching using treatment rank. Bars represent mean observed abstinence (from original trial) for individuals who received their model-predicted best, second-best, or worst treatment, over time. Error bars indicate standard errors."

fig_rank_wk4_supp <- d |> 
  mutate(week_cat = factor(week,
                           levels = c(4, 12, 26),
                           labels = c("Week 4", "Week 12", "Week 26")),
         tx_rank = factor(tx_rank, 
                           levels = c("first", "second", "third"),
                           labels = c("Best Treatment", "Second-best Treatment",
                                      "Worst Treatment"))) |> 
  group_by(week_cat, tx_rank) |> 
  summarize(mean_outcome = mean(outcome_rct_num),
            sd_outcome = sd(outcome_rct_num),
            n_group = n()) |> 
  mutate(se_outcome = (sd_outcome / sqrt(n_group))) |> 
  ggplot(aes(x = week_cat, y = mean_outcome, fill = tx_rank)) +
  geom_col(position = "dodge") +
  geom_errorbar(aes(ymin = mean_outcome - se_outcome, 
                    ymax = mean_outcome + se_outcome,
                    x = week_cat),
                position = position_dodge(width = 0.9),
                width = 0.5) +
  scale_y_continuous(limits = c(0, 0.45)) +
  labs(
    x = "Week On-Study",
    y = "Mean Abstinence Rate",
    fill = "Treatment Rank"
  )

fig_rank_wk4_supp
```
