---
title: "Make All Tables for Main Manuscript"
author: "Gaylen Fronk"
date: "`r lubridate::today()`"
number-sections: true
output: 
  html_document:
    toc: true 
    toc_depth: 4
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 4
editor_options: 
  chunk_output_type: console
html-table-processing: none
bibliography: ../references.bib
bibliographystyle: ../apa.csl
---

```{r, packages_script}
#| message: false
#| warning: false

# packages for script
library(tidyverse)

devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_eda.R?raw=true")
theme_set(theme_classic()) 
```

```{r, packages_workflow}
#| message: false
#| warning: false

# handle conflicts
options(conflicts.policy = "depends.ok")
```

```{r, absolute paths}

# absolute paths
switch (Sys.info()[['sysname']],
        # PC paths
        Windows = {
          path_data <- str_c("P:/studydata/match/processed_data/")
          path_models <- str_c("P:/studydata/match/models/pp_hybrid_wk4_outcome")},
        
        # IOS paths
        Darwin = {
          path_data <- str_c("/Volumes/private/studydata/match/processed_data/")
          path_models <- str_c("/Volumes/private/studydata/match/models/pp_hybrid_wk4_outcome")},
        
        # Linux paths
        Linux = {
          path_data <- str_c("~/mnt/private/studydata/match/processed_data/")
          path_models <- str_c("~/mnt/private/studydata/match/models/pp_hybrid_wk4_outcome")}
)
```

```{r defaults}

# chunk defaults

options(tibble.width = Inf)
options(tibble.print_max = Inf)
```

## Tables 1, 2, & 3 (Features & Sample Characteristics)

```{r}
d <- read_csv(file.path(path_data, "match_cln.csv"),
              show_col_types = FALSE)
```


### Table 1: Feature Categories

Set up data
```{r}
dt3 <- d |> select(-contains("outcome"), -subid)
```

Build table
```{r}
t1_feat_misc <- tibble(
  feat_cat = NA_character_,
  feat_name = c("Treatment", "Incarcerated", "Works Third Shift"),
  type = c("Categorical", "Dichotomous", "Dichotomous"),
  n_feat = c(1, 1, 1)
) |> 
  add_row(feat_cat = "Miscellaneous",
  feat_name = NA_character_,
  type = c(),
  n_feat = NA_real_,
  .before = 1)

t1_feat_psych_sud <- tibble(
  feat_cat = NA_character_,
  feat_name = c("Frequency of Drinking Alcohol", "Quantity of Alcohol", "Binge Drinking", "Short Inventory of Problems-2 (Revised) [@kilukShortInventoryProblems2013]", "Life Satisfaction", "Life Enjoyment", "Psychological Disorder Diagnoses", "Positive and Negative Affect Schedule [@crawfordPositiveNegativeAffect2004a]", "Snaith-Hamilton Pleasure Scale [@snaithScaleAssessmentHedonic1995a]", "Anxiety Severity Index-3 [@taylorRobustDimensionsAnxiety2007a]", "Distress Tolerance Scale [@simonsDistressToleranceScale2005a]", "Patient History Questionnaire-9 [@kroenkePHQ9ValidityBrief2001a]"),
  type = c("Ordinal", "Ordinal", "Ordinal", "Ordinal", "Ordinal", "Ordinal", "Dichotomous", "Ordinal", "Ordinal", "Ordinal", "Ordinal", "Ordinal"),
  n_feat = c(1, 1, 1, 16, 1, 1, 7, 8, 15, 22, 20, 11)
) |> 
  add_row(feat_cat = "Mental Health & Psychological Traits",
  feat_name = NA_character_,
  type = c(),
  n_feat = NA_real_,
  .before = 1)

t1_feat_smoking <- tibble(
  feat_cat = NA_character_,
  feat_name = c("Baseline Carbon Monoxide", "Carbon Monoxide Exposure", "Age of 1st Cigarette", "Age Became Daily Smoker", "Years Smoking", "Cigarettes Per Day (Heaviest)", "Use of Other Tobacco Products", "Number of Previous Quit Attempts", "Last Recent Quit Attempt", "Longest Quit Attempt", "Previous Quit Methods Used", "Cigarettes Per Day (Current)", "Motivation to Quit", "Self-Efficacy for Quitting in Next 30 Days", "Confidence to Quit", "Importance to Quit", "DSM5 Tobacco Use Disorder [@americanpsychiatricassociationDiagnosticStatisticalManual2013c]", "Fagerstrom Test of Nicotine Dependence [@heathertonFagerstromTestNicotine1991]", "Wisconsin Inventory of Smoking Dependence Motives-37 [@smithDevelopmentBriefWisconsin2010]", "Smoke Menthol Cigarettes", "Wisconsin Smoking Withdrawal Scale-2 [@smithRevisionWisconsinSmoking2021]"),
  type = c("Numeric", "Dichotomous", "Numeric", "Numeric", "Numeric", "Numeric", "Categorical (Unordered)", "Numeric", "Ordinal", "Ordinal", "Dichotomous", "Numeric", "Ordinal", "Ordinal", "Ordinal", "Ordinal", "Dichotomous", "Dichotomous, Ordinal", "Ordinal", "Categorical (Unordered)", "Ordinal"),
  n_feat = c(1, 1, 1, 1, 1, 1, 5, 1, 1, 1, 6, 1, 1, 1, 1, 1, 14, 6, 49, 1, 45)
) |> 
  add_row(feat_cat = "Smoking Use & History",
  feat_name = NA_character_,
  type = c(),
  n_feat = NA_real_,
  .before = 1)

t1_feat_medical <- tibble(
  feat_cat = NA_character_,
  feat_name = c("Diabetes Diagnosis", "Multidimensional Fatigue Inventory [@smetsMultidimensionalFatigueInventory1995]", "Berlin Sleep Questionnaire [@netzerUsingBerlinQuestionnaire1999]", "Body Mass Index", "Health-related Quality of Life Scale [@taylorMeasuringHealthyDays2000]", "Healthy Days Symptoms Module [@moriartyCDCStudiesCommunity1996]"),
  type = c("Dichotomous", "Ordinal", "Ordinal, Dichotomous", "Numeric", "Ordinal, Numeric", "Numeric"),
  n_feat = c(1, 26, 3, 1, 4, 5)
) |> 
  add_row(feat_cat = "Medical & Physical Health",
  feat_name = NA_character_,
  type = c(),
  n_feat = NA_real_,
  .before = 1)

t1_feat_social_environ <- tibble(
  feat_cat = NA_character_,
  feat_name = c("Spouse Smokes", "Live with Another Smoker", "People Close to You Who Smoke", "Ban on Smoking at Home", "Ban on Smoking at Work", "Time Around Other Smokers"),
  type = c("Categorical (unordered)", "Categorical (unordered)", "Categorical (unordered)", "Dichotomous", "Categorical (unordered)", "Ordinal"),
  n_feat = c(1, 1, 5, 1, 1, 2)
) |> 
  add_row(feat_cat = "Social & Environmental Characteristics",
  feat_name = NA_character_,
  type = c(),
  n_feat = NA_real_,
  .before = 1)

t1_feat_demo <- tibble(
  feat_cat = NA_character_,
  feat_name = c("Gender", "Age", "Race", "Marital Status", "Income", "Ethnicity", "Employment"),
  type = c("Categorical (unordered)", "Numeric", "Categorical (unordered)", "Categorical (unordered)", "Ordinal", "Categorical (Unordered)", "Categorical (unordered)"),
  n_feat = c(1, 1, 1, 1, 1, 1, 1)
) |> 
  add_row(feat_cat = "Demographic Characteristics",
  feat_name = NA_character_,
  type = c(),
  n_feat = NA_real_,
  .before = 1)

t1_feat <- bind_rows(t1_feat_demo, t1_feat_smoking) |> 
  bind_rows(t1_feat_social_environ) |> 
  bind_rows(t1_feat_psych_sud) |> 
  bind_rows(t1_feat_medical) |> 
  bind_rows(t1_feat_misc) |> 
  rename(Category = feat_cat,
         `Feature Name` = feat_name,
         Type = type,
         `Number of Features` = n_feat)
```

Table 1
```{r tbl-featureshtml}
#| label: tbl-featureshtml
#| tbl-cap: "Individual Differences"

t1_feat |>
  mutate(across(where(is.numeric), as.character)) |>
  mutate(across(everything(), ~replace_na(., "")))  |>
  knitr::kable() 

```

Table 1, formatted for Typst/PDF output
```{r tbl-featurespdf}
#| label: tbl-featurespdf
#| tbl-cap: "Individual Differences"

t1_feat_demo |>
  mutate(feat_cat = " ") |>
  rename(`Demographic Characteristics` = feat_cat,
         `Feature Name` = feat_name,
         Type = type,
         `# Features` = n_feat) |>
tail(-1) |>
  knitr::kable(align = "lllc") 

```


```{r featurespdf2}
#| label: featurespdf2

t1_feat_smoking |>
  mutate(feat_cat = " ") |>
  rename(`Smoking Use/History` = feat_cat,
         `Feature Name` = feat_name,
         Type = type,
         `# Features` = n_feat) |>
tail(-1) |>
  knitr::kable(align = "lllc") 

```


```{r featurespdf3}
#| label: featurespdf3

t1_feat_social_environ |>
  mutate(feat_cat = " ") |>
  rename(`Social & Environmental Characteristics` = feat_cat,
         `Feature Name` = feat_name,
         Type = type,
         `# Features` = n_feat) |>
tail(-1) |>
  knitr::kable(align = "lllc") 

t1_feat_psych_sud |>
  mutate(feat_cat = " ") |>
  rename(`Mental Health & Psychological Traits` = feat_cat,
         `Feature Name` = feat_name,
         Type = type,
         `# Features` = n_feat) |>
tail(-1) |>
  knitr::kable(align = "lllc") 

```


```{r featurespdf4}
#| label: featurespdf4

t1_feat_medical |>
  mutate(feat_cat = " ") |>
  rename(`Medical & Physical Health` = feat_cat,
         `Feature Name` = feat_name,
         Type = type,
         `# Features` = n_feat) |>
tail(-1) |>
  knitr::kable(align = "lllc") 

t1_feat_misc |>
  mutate(feat_cat = " ") |>
  rename(`Miscellaneous Features` = feat_cat,
         `Feature Name` = feat_name,
         Type = type,
         `# Features` = n_feat) |>
tail(-1) |>
   knitr::kable(align = "lllc") 
```

### Table 2: Demographic Characteristics

Select down data
```{r}
dt2 <- d |> 
  select(employment, ends_with("_ehr"), income_order) |> 
  select(-bmi_ehr, -diabetic_ehr) |> 
  glimpse()
```

Build table
```{r}
tab_race <- dt2 |> 
  tab(race_ehr) |> 
  mutate(race_ehr = case_match(
    race_ehr,
    "asian" ~ "   Asian",
    "black_african_american" ~ "   Black/African American",
    "multiracial" ~ "   Multiracial",
    "native_american_alaska_native" ~ "   Native American/Alaska Native",
    "other" ~ "   Other",
    "white" ~ "   White"
  )) |> 
  rename(var_name = race_ehr) |> 
  mutate(n_prop = str_c(n, " (", round((prop * 100), digits = 2), "%)")) |> 
  select(var_name, n_prop) |> 
  add_row(var_name = "Race",
          n_prop = "",
          .before = 1)

tab_ethnicity <- dt2 |> 
  tab(ethnicity_ehr) |> 
  mutate(ethnicity_ehr = case_match(
    ethnicity_ehr,
    "hispanic_latinx" ~ "   Hispanic or Latino/a",
    "non_hispanic" ~ "   Non-Hispanic"
  )) |> 
  rename(var_name = ethnicity_ehr) |> 
  mutate(n_prop = str_c(n, " (", round((prop * 100), digits = 2), "%)")) |> 
  select(var_name, n_prop) |> 
  add_row(var_name = "Ethnicity",
          n_prop = "",
          .before = 1)

tab_gender <- dt2 |> 
  tab(gender_ehr) |> 
  mutate(gender_ehr = case_match(
    gender_ehr,
    "female" ~ "   Female",
    "male" ~ "   Male"
  )) |> 
  rename(var_name = gender_ehr) |> 
  mutate(n_prop = str_c(n, " (", round((prop * 100), digits = 2), "%)")) |> 
  select(var_name, n_prop) |> 
  add_row(var_name = "Gender",
          n_prop = "",
          .before = 1)

tab_employment <- dt2 |> 
  tab(employment) |> 
  mutate(employment = case_match(
    employment,
    "employed_full_time" ~ "   Employed (full-time)",
    "employed_part_time" ~ "   Employed (part-time)",
    "unemployed" ~ "   Unemployed"
  )) |> 
  rename(var_name = employment) |> 
  mutate(n_prop = str_c(n, " (", round((prop * 100), digits = 2), "%)")) |> 
  select(var_name, n_prop) |> 
  add_row(var_name = "Employment",
          n_prop = "",
          .before = 1)

tab_marital <- dt2 |> 
  tab(marital_status_ehr) |> 
  mutate(marital_status_ehr = case_match(
    marital_status_ehr,
    "divorced" ~ "   Divorced",
    "live_with_domestic_partner" ~ "   Living with a domestic partner",
    "married" ~ "   Married",
    "separated" ~ "   Separated",
    "widowed" ~ "   Widowed",
    "never_married" ~ "   Never married",
    NA ~ "   Did not respond"
  )) |> 
  rename(var_name = marital_status_ehr) |> 
  mutate(n_prop = str_c(n, " (", round((prop * 100), digits = 2), "%)")) |> 
  select(var_name, n_prop) |> 
  add_row(var_name = "Marital Status",
          n_prop = "",
          .before = 1)

tab_income <- dt2 |> 
  tab(income_order) |> 
  mutate(income_order = case_match(
    income_order,
    "01_less_than_10000" ~ "   < $10,000",
    "02_btwn_10000_19999" ~ "   $10,000 - $19,999",
    "03_btwn_20000_24999" ~ "   $20,000 - $24,999",
    "04_btwn_25000_34999" ~ "   $25,000 - $34,999",
    "05_btwn_35000_49999" ~ "   $35,000 - $49,999",
    "06_btwn_50000_74999" ~ "   $50,000 - $74,999",
    "07_more_than_75000" ~ "   $75,000+",
    NA ~ "   Did not respond"
  )) |> 
  rename(var_name = income_order) |> 
  mutate(n_prop = str_c(n, " (", round((prop * 100), digits = 2), "%)")) |> 
  select(var_name, n_prop) |> 
  add_row(var_name = "Income",
          n_prop = "",
          .before = 1)

tab_age <- tibble(
  var_name = "Age",
  n_prop = "",
  `Mean (SD)` = str_c(round(mean(dt2$age_ehr), digits = 2),
                      " (", round(sd(dt2$age_ehr), digits = 2), ")")
)

t2_demo <- bind_rows(tab_gender, tab_race) |> 
  bind_rows(tab_ethnicity) |> 
  bind_rows(tab_marital) |> 
  bind_rows(tab_employment) |> 
  bind_rows(tab_income) |> 
  mutate(`Mean (SD)` = "") |> 
  add_row(tab_age, .before = 1) |> 
  rename(Characteristic = var_name,
         `N (%)` = n_prop)
  
```

Table 2
```{r}
#| label: tbl-demohtml
#| tbl-cap: Demographic Characteristics

t2_demo |>
  knitr::kable()

```

```{r}
#| label: tbl-demopdf
#| tbl-cap: Demographic Characteristics

t2_demo |>
  head(-9) |>
  knitr::kable()

```

```{r}
#| label: demopdf

t2_demo |>
  tail(9) |>
  knitr::kable()

```


### Table 3: Smoking-related ID Characteristics

Select down data
```{r}
dt3 <- d |> 
  select(age_1st_cig, age_daily_smoker, years_smoke, cigs_per_day_heaviest,
         tot_quit_attempt, cigs_per_day, dsm5_total_scale, ftnd_1_item_order,
         wisdm37_total_scale, wsws_total_scale)
```

Build table
```{r}
tab_ftnd_1 <- dt3 |> 
  tab(ftnd_1_item_order) |> 
  mutate(ftnd_1_item_order = case_match(
    ftnd_1_item_order,
    "03_within_5_min" ~ "   Within 5 minutes",
    "02_6_to_30_min" ~ "   6 - 30 minutes",
    "01_31_to_60_min" ~ "   31 - 60 minutes",
    "00_after_60_min" ~ "   After 60 minutes",
    NA ~ "   Did not respond"
  )) |> 
  rename(var_name = ftnd_1_item_order) |> 
  mutate(n_prop = str_c(n, " (", round((prop * 100), digits = 2), "%)")) |> 
  select(var_name, n_prop) |> 
  add_row(var_name = "Time to first cigarette upon waking",
          n_prop = "",
          .before = 1) |> 
  rename(`N (%)` = n_prop, Characteristic = var_name) |> 
  mutate(`Mean (SD)` = "")

tab3_means <- dt3 |> 
  select(-ftnd_1_item_order) |> 
  summarize(across(everything(), ~ mean(.x, na.rm = TRUE))) |> 
  pivot_longer(
    cols = everything(),
    names_to = "var_name",
    values_to = "means"
  ) |> 
  mutate(means = round(means, digits = 2))

tab3_sds <- dt3 |> 
  select(-ftnd_1_item_order) |> 
  summarize(across(everything(), ~ sd(.x, na.rm = TRUE))) |> 
  pivot_longer(
    cols = everything(),
    names_to = "var_name",
    values_to = "sds"
  ) |> 
  mutate(sds = round(sds, digits = 2))

t3_idchars <- full_join(tab3_means, tab3_sds, by = "var_name") |> 
  mutate(var_name = case_match(
    var_name,
    "age_1st_cig" ~ "Age of first cigarette",
    "age_daily_smoker" ~ "Age became daily smoker",
    "years_smoke" ~ "Years smoking",
    "cigs_per_day" ~ "Cigarettes per day (current)",
    "cigs_per_day_heaviest" ~ "Cigarettes per day (heaviest)",
    "tot_quit_attempt" ~ "Number of previous quit attempts",
    "dsm5_total_scale" ~ "Number of DSM5 tobacco use disorder symptoms",
    "wisdm37_total_scale" ~ "WISDM37 Total Score",
    "wsws_total_scale" ~ "WSWS Total Score"
  )) |> 
  mutate(`Mean (SD)` = str_c(means, " (", sds, ")"),
         `N (%)` = "") |> 
  rename(Characteristic = var_name) |> 
  select(-means, -sds) |> 
  full_join(tab_ftnd_1,
            by = join_by(Characteristic, `Mean (SD)`, `N (%)`))
  
```

Table 3
```{r}
#| label: tbl-smoking-chars
#| tbl-cap: Smoking-related ID Characteristics

t3_idchars |>
  knitr::kable()


```

## Tables 4, 5, & 6 (Treatment Interactions)

Read in data
```{r}
model_tidy <- read_csv(file.path(path_models, "model_best_tidy_v6.csv"),
                       show_col_types = FALSE) |> 
  glimpse()
```

Make overall starting table
```{r}
retained_vars_tx <- model_tidy |> 
  mutate(estimate = estimate * -1) |> 
  filter(abs(estimate) > 0) |> 
  select(-penalty) |> 
  filter(str_detect(term, "treatment_")) |> 
  arrange(desc(abs(estimate)))

print_kbl(retained_vars_tx, digits = 4)
```

### Table 4: Varenicline Interactions

Make table
```{r}
varenicline <- retained_vars_tx |> 
  filter(str_detect(term, "varenicline")) |> 
  mutate(term = str_replace(term, "treatment_varenicline_x_", "")) |> 
  arrange(desc(abs(estimate))) |> 
  mutate(direction = if_else(estimate > 0, "+", "-")) |> 
  mutate(Feature = c(
    "Lives alone or only with a partner",
    "Worries they will choke to death when their throat feels tight (ASI-3 Item 15)",
    "Does not think they do a lot in a day (MFI Item 6)",
    "Income (Increasing income bracket)",
    "Has a spouse/partner who does not smoke",
    "No close relative who smokes",
    "Kept smoking despite relationship problems (DSM-5 Item 3)",
    "Has never tried cigars",
    "Has a depression diagnosis",
    "Does not live with another smoker",
    "Smoking helps them think better (WISDM-37 Item 32)",
    "Feels tired, fatigued, or not up to par (Berlin Item 7)",
    "Has tried cigars but has never used regularly",
    "A lot of friends or family members smoke (WISDM-37 Item 27)",
    "Has a close relative who smokes",
    "Bothered by constipation (WSWS-2 Item 'Constipation')",
    "Smoked to get rid of withdrawal feelings (DSM-5 Item 7)",
    "Tried nicotine gum during previous quit attempt",
    "Has a spouse/partner who smokes",
    "Has not tried nicotine gum during a previous quit attempt",
    "Feeling hopeless or discouraged (WSWS-2 Item 19)",
    "Most of the people they spend time with are smokers (WISDM-37 Item 18)",
    "Doesn't live with a spouse/partner",
    "Smoking only allowed in some parts of workplace"
  )) |> 
  select(Feature, `Effect Direction` = direction)

```

```{r}
#| label: tbl-int-varenicline
#| tbl-cap-location: bottom
#| tbl-cap: "Retained features that interact with varenicline. Direction of effect indicates whether that feature increases or decreases the probability of abstinence when using varenicline (compared to C-NRT and nicotine patch)."

varenicline |> 
  knitr::kable(align = "lc")

```

### Table 5: Combo-NRT Interactions

Make table
```{r}
combo_nrt <- retained_vars_tx |> 
  filter(str_detect(term, "combo_nrt")) |> 
  mutate(term = str_replace(term, "treatment_combo_nrt_x_", "")) |> 
  arrange(desc(abs(estimate))) |> 
  mutate(direction = if_else(estimate > 0, "+", "-")) |> 
  mutate(Feature = c(
    "Lives alone or only with a partner",
    "Identifies as Black or African American",
    "Sometimes feels like cigarattes are their best friends (WISDM-37 Item 22)",
    "Has never smoked menthol cigarettes",
    "Disagrees that there is nothing worse than feeling distressed or upset (DTS Item 5)",
    "Most of their friends and acquaintances smoke (WISDM-37 Item 30)",
    "Divorced",
    "Does not endorse withdrawal symptoms (DSM-5 Item 6)",
    "Worries they are going crazy when they cannot keep their mind on a task (ASI-3 Item 2)",
    "Identifies as White",
    "Married",
    "Has spent a lot of time obtaining, smoking, or recovering from cigarettes (DSM-5 Item 10)",
    "Currently smokes menthol cigarettes",
    "Endorses weight control as a major reason they smoke (WISDM-37 Item 19)",
    "Does not feel it takes a lot of effort to concentrate (MFI Item 13)",
    "Does not smoke when so ill they are in bed most of the day (FTND Item 6)",
    "Lives with another smoker",
    "Satisfied with their life as a whole",
    "Bothered by constipation (WSWS-2 Item 'Constipation')",
    "Has tried a pipe but has never used regularly",
    "Amount of time since most recent quit attempt",
    "Can handle feeling distressed or upset (DTS Item 3)",
    "Smokes when so ill they are in bed most of the day (FTND Item 6)",
    "Has never been diagnosed with sleep apnea (Berlin Item 9)",
    "Bothered by waking frequently during the night (WSWS-2 Item 5)",
    "Bothered by having urges to smoke (WSWS-2 Item 9)",
    "More days in past month feeling worried, tense, or anxious (HDSM Item `Tense`)",
    "Bothered by wanting to smoke (WSWS-2 Item 'Want Cigarette')",
    "Kept smoking knowing it was causing or worsening a health problem (DSM-5 Item 12)",
    "Does not find feeling distressed or upset to be unbearable (DTS Item 1)",
    "Tried Zyban (bupropion) during previous attempt"
  )) |> 
  select(Feature, `Effect Direction` = direction)
```

```{r}
#| label: tbl-int-cnrt
#| tbl-cap-location: bottom
#| tbl-cap: "Retained features that interact with combination nicotine replacement therapy (C-NRT). Direction of effect indicates whether that feature increases or decreases the probability of abstinence when using C-NRT (compared to varenicline and nicotine patch)."

combo_nrt |> 
  knitr::kable(align = "lc")

```

### Table 6: Patch Interactions

Make table
```{r}
patch <- retained_vars_tx |> 
  filter(str_detect(term, "patch")) |> 
  mutate(term = str_replace(term, "treatment_patch_x_", "")) |> 
  arrange(desc(abs(estimate))) |> 
  mutate(direction = if_else(estimate > 0, "+", "-")) |> 
  mutate(Feature = c(
    "Has never tried cigars",
    "No close friend who smokes",
    "Has tried cigars but has never used regularly",
    "Has a close co-worker who smokes",
    "No smoking allowed anywhere at their workplace",
    "Bothered by craving a cigarette (WSWS-2 Item `Crave`)",
    "Endorses feeling alone without their cigarettes (WISDM-37 Item 26)",
    "Has a close friend who smokes",
    "Has never been married",
    "Identifies as female",
    "Finds it difficult to refrain from smoking where it is forbidden (FTND Item 2)",
    "Endorses feeling that cigarettes rule their life (WISDM-37 Item 16)",
    "Identifies as White",
    "Does not work outside the home",
    "Bothered by coughing (WSWS-2 Item `Coughing`)",
    "Does not have a lot of plans (MFI Item 15)",
    "Identifies as male",
    "Bothered by having trouble getting cigarettes of their mind (WSWS-2 Item 26)",
    "Greater total number of previous quit attempts"
  )) |> 
  select(Feature, `Effect Direction` = direction)
```

```{r}
#| label: tbl-int-patch
#| tbl-cap-location: bottom
#| tbl-cap: "Retained features that interact with nicotine patch. Direction of effect indicates whether that feature increases or decreases the probability of abstinence when using nicotine patch (compared to varenicline and C-NRT)."

patch |> 
  knitr::kable(align = "lc")

```