---
title: "Make Figure 1: Model Performance (3-Panel)"
author: "Gaylen Fronk"
date: "`r lubridate::today()`"
number-sections: true
output: 
  html_document:
    toc: true 
    toc_depth: 4
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 4
editor_options: 
  chunk_output_type: console
---

## Setup

```{r set_values}
version <- "v5"
cv <- "nested"
y_col_name <- "pp_hybrid_wk4_outcome"
```

```{r, packages_script}
#| message: false
#| warning: false

# packages for script
library(tidyverse)
library(tidymodels)
library(tidyposterior)
library(probably)
library(patchwork)

devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_eda.R?raw=true")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/chtc/static_files/fun_chtc.R?raw=true")
theme_set(theme_classic()) 
```

```{r, packages_workflow}
#| message: false
#| warning: false

# handle conflicts
options(conflicts.policy = "depends.ok")
```

```{r, absolute_paths}
switch (Sys.info()[['sysname']],
        # PC paths
        Windows = {
          path_models <- str_c("P:/studydata/match/models/", 
                               y_col_name)
        },
        
        # IOS paths
        Darwin = {
          path_models <- str_c("/Volumes/private/studydata/match/models/",
                               y_col_name)
        },
        
        # Linux paths
        Linux = {
          path_models <- str_c("~/mnt/private/studydata/match/models/",
                               y_col_name)
        }
)
```

```{r defaults}

# chunk defaults

options(tibble.width = Inf)
options(tibble.print_max = Inf)
```

## Make Panels

### Panel A: ROC Curve

Read in outer loop predictions
```{r}
preds_out <- read_rds(file.path(path_models, 
                                str_c("outer_preds_", version, 
                                      "_", cv, ".rds"))) |> 
  glimpse()
```

Set up data
```{r}
roc_data <- preds_out %>%
  roc_curve(prob_raw, truth = label) |> 
  glimpse()
```

Create plot
```{r}
#| label: fig-roc-single
fig_roc <- roc_data %>%
  ggplot(aes(x = 1 - specificity, y = sensitivity, color = .threshold)) +
  geom_path(linewidth = 2) +
  geom_abline(lty = 3) +
  coord_fixed(xlim = c(0, 1), ylim = c(0, 1)) +
  labs(x = "False Positive Rate",
       y = "True Positive Rate",
       color = "Threshold") +
  scale_x_continuous(breaks = seq(0,1,.25),
                     labels = sprintf("%.2f", seq(1,0,-.25))) +
  scale_color_gradient(low="blue", high="red") +
  theme(axis.text = element_text(size = rel(1.25)),
        axis.title = element_text(size = rel(1.5)))

fig_roc
```

### Panel B: Prediction Calibration

Read in final model predictions (LOOCV)
```{r}
d_cal <- read_csv(file.path(path_models,
                            str_c("aim_2_", version, "_", y_col_name, ".csv")),
                  show_col_types = FALSE) |> 
  glimpse()
```

Set up data
```{r}
d_cal <- d_cal |> 
  mutate(.pred_abstinent = case_when(
    tx_rct == "patch" ~ prob_patch,
    tx_rct == "combo_nrt" ~ prob_combo_nrt,
    tx_rct == "varenicline" ~ prob_varenicline,
    TRUE ~ NA_real_
  )) |> 
  select(outcome_rct = outcome_rct_wk4, .pred_abstinent) |> 
  mutate(outcome_rct = factor(outcome_rct,
                              levels = c("abstinent", "smoking"))) |> 
  glimpse()
```

Create breaks plot
```{r}
#| label: fig-cal-breaks
fig_cal_breaks <- d_cal |> 
  cal_plot_breaks(truth = outcome_rct, 
                  estimate = .pred_abstinent)

fig_cal_breaks
```

Create windowed plot
```{r}
#| label: fig-cal-window
fig_cal_window <- d_cal |> 
  cal_plot_windowed(truth = outcome_rct, 
                    estimate = .pred_abstinent)

fig_cal_window
```

Created binned plot
```{r}
#| label: fig-cal-bins
bin_width = 0.10
fig_cal_bins <- d_cal |> 
  rename(prob_raw = .pred_abstinent) |> 
  mutate(bins = cut(prob_raw, breaks = seq(0, 1, bin_width)),
         abstinent = if_else(outcome_rct == "abstinent", 1, 0)) |> 
  group_by(bins)  |> 
  summarize(mean_abstinent = mean(abstinent),
            .groups = "drop") |>
  mutate(bins = as.numeric(bins),
         midpoints = bin_width/2 + bin_width * (bins - 1))  |> 
  ggplot(data = _, aes(x = midpoints, y = mean_abstinent)) +
    geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
    geom_line() +
    geom_point() +
    xlab("Predicted Abstinence Probability (bin mid-point)") +
    ylab("Observed Abstinence Probability") +
    scale_x_continuous(breaks = seq(0, 1, bin_width),
                       limits = c(0, 1)) +
    scale_y_continuous(breaks = seq(0, 1, bin_width),
                       limits = c(0, 1)) +
  theme(axis.text = element_text(size = rel(1.25)),
        axis.title = element_text(size = rel(1.5)))

fig_cal_bins
```

### Panel C: Posterior Probability Distribution

Read in posterior probabilities
```{r}
#| eval: false
pp <- read_rds(file.path(path_models, 
                         str_c("posteriors_", version, "_nested.rds"))) 
```

Set up data
```{r}
#| eval: false
pp_tidy <- pp %>% 
  tidy(seed = 123) 

ci_week4 <- pp_tidy |> 
  summary() |> 
  mutate(model = factor(model, 
                        levels = "roc_auc",
                        labels = "Week 4"),
         y = 1000) 

```

Create plot
```{r}
#| eval: false
#| label: fig-pp-wk4
fig_pp <- pp_tidy |> 
  mutate(model = factor(model, 
                        levels = "roc_auc",
                        labels = "Week 4")) |> 
  ggplot() + 
  geom_histogram(aes(x = posterior, fill = model), 
                 color = "black", alpha = .4, 
                 bins = 30) +
  geom_segment(mapping = aes(y = y + 100, yend = y - 100, 
                             x = mean, xend = mean,
                             color = model),
               data = ci_week4) +
  geom_segment(mapping = aes(y = y, yend = y, 
                             x = lower, xend = upper, 
                             color = model),
               data = ci_week4) +
  scale_y_continuous("Posterior Probability", 
                     breaks = c(0, 250, 500, 750)) +
  xlab("Area Under ROC Curve") +
  theme(legend.position = "none")

fig_pp
```

## Combine into single figure

```{r}
#| eval: false
patchwork <- fig_roc + fig_cal_bins

patchwork + plot_annotation(
  title = "Model Performance",
  tab_levels = "A"
) +
  plot_layout(ncol = 1)
```

